### 2025-06-19 reports/paper/cover_letter.tex

\documentclass{article}

\usepackage{charter} % Use the Charter font for the document text
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr,graphicx}

\fancypagestyle{firstpage}{%
  \fancyhf{}% Clear header/footer
  \renewcommand{\headrulewidth}{0pt}%
  \renewcommand{\footrulewidth}{1pt}%
  % Add more detail here if needed
}
\fancypagestyle{otherpages}{%
  \fancyhf{}% Clear header/footer
  \renewcommand{\headrulewidth}{1pt}%
  \renewcommand{\footrulewidth}{1pt}%
  % Add more detail here if needed
}

\AtBeginDocument{\thispagestyle{firstpage}}
\pagestyle{otherpages}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex}

\begin{document}

\vspace*{\dimexpr-\headsep-\headheight-1pt}

%\includegraphics[width=2.1in]{logo.png}

\rule{\linewidth}{1pt}

\bigskip\bigskip

\hfill
\begin{tabular}{ l @{} }
  \today \\[12pt] % Date
\end{tabular}

\bigskip

\begin{tabular}{ @{} l }
  Viktor Andonovikj \\ % Addressee of the letter above the to address
  Young researcher \\ % To address
  Jožef Stefan Institute \\
  Jamova cesta 39 \\
  1000 Ljubljana\\
  Slovenia
\end{tabular}

\bigskip

Dear Editor,

\bigskip

On behalf of my co-authors, I am submitting our paper titled ``Variational oblique predictive clustering trees''.

The paper presents a novel model that integrates Bayesian inference into oblique predictive clustering trees to enhance predictive performance, interpretability, and uncertainty quantification within a single-tree framework. Our results show that the proposed model achieves performance competitive with ensemble methods while retaining the transparency of interpretable models.

We believe that the paper provides a valuable contribution to the field, and we hope that it will be of interest to the readership of Expert Systems with Applications.

Thank you for considering our submission.

\bigskip

Sincerely yours,

\vspace{50pt}

Viktor Andonovikj

\end{document}