### 2025-06-19 reports/paper/credit_author_statement.tex

\documentclass{article}
\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\begin{document}
\begin{description}
\item[Viktor Andonovikj] Code writing, Manuscript writing, Data analysis, Modelling

\item[Sa\v{s}o D\v{z}eroski] Supervision, Concept validation, Reviewing and editing

\item[Biljana Mileva Boshkoska] Supervision, Conceptualisation, Data analysis, Reviewing and editing

\item[Pavle Bo\v{s}koski]  Supervision, Conceptualisation, Data provision, Data analysis, Reviewing and editing

\end{description}

\end{document}