### 2025-06-19 reports/paper/highlights_vspyct.tex

\documentclass{article}

\begin{document}
\begin{itemize}
\item Introduces VSPYCT, a variational Bayesian oblique decision tree for structured output prediction.
\item Captures parameter uncertainty through variational inference in each tree split.
\item Matches or outperforms SPYCT ensembles on classification and multi-target regression tasks.
\item Provides feature importance scores and visual interpretability within a single-tree framework.
\item Demonstrates robustness to spurious features and performs well across varied dataset properties.
\end{itemize}

\end{document}z